---
title: "Lab 8"
author: "PUBPOL 599"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this lab, you will create a dashboard relating to some mock experiment data that is created below. Someone wanted to see whether their "mindfulness", as rated by a daily app, would increase if they started meditating for ten minutes. They tracked their mindfulness for 28 days, and meditated the second set of two weeks. They also collected data on the number of steps they got, and whether they had any tests - they thought these two things might also impact their mindfulness rating.

In a one tab flex dashboard, create visualizations to display the data generated below and to assess the value of the experiment.

*Note* The data is randomly generated, so it is very likely there will not be any affect. That is okay in this assignment and your own, just display the data as best you can to indicate what happened in either direction. 

```{r, echo=T}
set.seed(123)
day <- 1:28
meditate <-  ifelse(day > 14, 1, 0)
mindful <- runif(28, 70, 100)
steps <- runif(28, 8000, 13500)
tests <- rbinom(28, 1, 0.12)
experiment = as.data.frame(cbind(day, meditate, mindful, steps, tests))

head(experiment)
```

Step 1. Generate dashboard

Step 2. Create at least three visualizations to show what did or didn't affect mindfulness

Step 3. Include evidence for the experiments success or failure
