---
title: "Evaluating Experiments CLass Example"
author: "PUBPOL 599"
date: "5/19/2019"
output: html_document
---

### Generating Data from Eric's Sleep experiment

```{r}
library(ggplot2)
library(pander)
```

On day 15, Eric committed to get into bed at 10 every night, and tracked his nightly sleep. Did that have an effect?

We have data based on the day, the minutes of sleep, whether the period was during the treatment or not, and whether the treatment was actually applied
```{r}
day <- 1:33
minutes <- c(431, 435, 432, 428, 436, 454, 429, 424, 410, 402, 403, 462,398, 430,
           443, 450, 462, 461, 457, 463, 471, 452, 439, 441, 440, 441, 438, 456, 448, 451, 439, 463, 452 )
treat <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,
          1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)
by10 <- c(0,0,0,1,0,0,0,0,0,1,0,0,1,0,
           1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)
start <- c(41, 59, 45, 43,51, 77,54, 52,77,45,24,50,
           27,49, 62, 65, 70, 42,41, 23, 35, 31, 40,
           40, 38,41, 32,22,31,28, 35,36, 51)

```

Let's combine those columns, and divide them into a pre and post data set for later use.

```{r}
 sleep <- as.data.frame(cbind(day, minutes, treat, by10, start))
sleep.pre <- subset(sleep,treat==0)
sleep.post <- subset(sleep,treat==1)
```

### Plotting

We can begin with a basic chart laying the data out over time with few bells and whistles

```{r}
ggplot(data = sleep, aes(x = day, y = minutes)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(300, 600)) + 
  theme_bw()
```

Let's add an indicator of the treatment so that we can see where we expect the data to change with geom_vline

```{r}
ggplot(data = sleep, aes(x = day, y = minutes)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(300, 600)) +
  geom_vline(xintercept=14.5,linetype="dashed") +
  geom_text(x = 20, y= 600, label="Treatment: in bed by 10") +
  theme_bw()
```

Great, that helps us identify where a change, but a trend line would help more. We can add that with geom_smooth.

```{r}
ggplot(data = sleep, aes(x = day, y = minutes)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(300, 600)) +
  geom_vline(xintercept=14.5,linetype="dashed") +
  geom_text(x = 20, y= 600, label="Treatment: in bed by 10") +
  geom_smooth(method="lm", se=FALSE, color="red") +
  theme_bw()
```

That's useful, and it shows an increase in sleep, but two lines would add additional information. We need to call geom_smooth twice then, using a subset of the data.

```{r}
ggplot(data = sleep, aes(x = day, y = minutes)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(300, 600)) +
  geom_vline(xintercept=14.5,linetype="dashed") +
  geom_text(x = 20, y= 600, label="Treatment: in bed by 10") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==0),se=FALSE,linetype="dotted") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==1),se=FALSE,linetype="dotted") +
  theme_bw()
```

What impact does that plot show?

Let's add a little bit more information to help someone evaluate the experiment. We can add the means for the pre and post period to clarify what happened. 

```{r}
ggplot(data = sleep, aes(x = day, y = minutes)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(300, 600)) + 
  geom_vline(xintercept=14.5,linetype="dashed") +
  geom_text(x = 20, y= 600, label="Treatment: in bed by 10") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==0),se=FALSE,linetype="dotted") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==1),se=FALSE,linetype="dotted") +
  geom_text(x=8, y=550, label="pre: 427 minutes") +
  geom_text(x=24, y=550, label="post: 451 minutes") +
  theme_bw()
```

Finally, let's add a title and clear axis labels to make sure someone could evaluate 

```{r}
ggplot(data = sleep, aes(x = day, y = minutes)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(300, 600)) + 
  geom_vline(xintercept=14.5,linetype="dashed") +
  geom_text(x = 20, y= 600, label="Treatment: in bed by 10") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==0),se=FALSE,linetype="dotted") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==1),se=FALSE,linetype="dotted") +
  geom_text(x=8, y=550, label="pre: 427 minutes") +
  geom_text(x=24, y=550, label="post: 451 minutes") +
  ggtitle("Minutes of Sleep Per Night") + 
  labs(y= "Minutes of Sleep", x = "Days") + 
  theme_bw()
```


A scatter plot or line graph are the most common ways to display such data, but they aren't the only way that's possible. Visualization has a lot of tools, try to find the best way to represent the data.

```{r}
ggplot(data = sleep, aes(x = day, y=minutes)) + 
  geom_bar(stat="identity") + 
  expand_limits(y=c(0, 600)) +
  geom_vline(xintercept=14.5, color="red",size=2,linetype="dotted")+
  geom_text(x = 20, y= 600, label="Treatment: in bed by 10", color="red") +
  geom_text(x=8, y=550, label="pre: 427 minutes") +
  geom_text(x=24, y=550, label="post: 451 minutes") +
  ggtitle("Minutes of Sleep Per Night") + 
  labs(y= "Minutes of Sleep", x = "Days") + 
  theme_bw()
```

So there is a difference. Is that different large? Is it significant? Is it meaningful?

We can use a t-test to assess significance

```{r}
t.test(sleep$minutes~sleep$treat)
```

brief aside, let's use pander to make that table look more attractive

```{r}
pander(t.test(sleep$minutes~sleep$treat))
```

Let's add that information to the plot by adding another labs() at the end

```{r}
ggplot(data = sleep, aes(x = day, y=minutes)) + 
  geom_bar(stat="identity") + 
  expand_limits(y=c(0, 600)) +
  geom_vline(xintercept=14.5, color="red",size=2,linetype="dotted")+
  geom_text(x = 18, y= 600, label="Treatment: in bed by 10", color="red") +
  geom_text(x=8, y=550, label="pre: 427 minutes") +
  geom_text(x=24, y=550, label="post: 451 minutes") +
  ggtitle("Minutes of Sleep Per Night") + 
  labs(y= "Minutes of Sleep", x = "Days") + 
  theme_bw() +
  labs(caption = "Difference in means significant (p=.0003")
```

So that is some evidence of an impact, but are there alternative theories? Let's see what happened with the time getting to sleep, rather than time into bed.

```{r}
ggplot(data = sleep, aes(x = day, y = start)) + 
  geom_point(color = "blue",shape=16, size=3) + 
  expand_limits(y=c(0, 120)) + 
  geom_vline(xintercept=14.5,linetype="dashed") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==0),se=FALSE,linetype="dotted") +
  geom_smooth(method = "lm", color="red", fullrange=F, data=subset(sleep,treat==1),se=FALSE,linetype="dotted") +
  ggtitle("Time Fell Asleep after 10") + 
  labs(y= "Minutes of After 10", x = "Days") + 
  theme_bw()

```

And what about the t-test for that change?


```{r}
t.test(sleep$start~sleep$treat)

```
